---
title: "Random walk simulations based on PCA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

`````{r}
rm(list=ls())
``````

`````{r, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(data.table)
library(stringr)
library(ggplot2)
library(zoo)
library(OneR)
library(fdasrvf)
library(ape)
library(ade4)
library(adegenet)
library(mvtnorm)
`````

#### data
This looks at the Mantamados (modern) pots.  The are three classes of pots: small, medium and large and several pots of each. The data are in long format.  The vases are not registered, but they have been scaled to have the same area.

`````{r}
a<-read.table("Mantamados_vases_outlines_norman.txt", header=FALSE, colClasses=c(V1="character"), fill=TRUE)
`````````
clean up sort out this data for use in R
`````{4}
a$datatype<-rep(c(rep("nondata",3), 1:200, "vasename"), 25)
a1<-a%>%
filter(datatype!="nondata")
vaseids<-a%>%
filter(datatype=="vasename")
vaseids1<-vaseids$V1
vaseids2<-gsub("IMAGE=", "", vaseids1)
a1$vase_id<-rep(vaseids2, each=201)
a2<-a1%>%
filter(datatype!="vasename")
a2$datatype<-NULL
names(a2)<-c("x", "y", "id")
`````

```{r}
a2$x<-as.numeric(as.character(a2$x))
a2$y<-as.numeric(as.character(a2$y))
```````
plot to look at what we have
`````{r}
ggplot(data=a2, aes(x=x, y=y, group=id))+
#geom_point(size=0.05)+
geom_polygon(fill=FALSE, size=0.05, colour="black")+
theme_classic()+
#guides(colour=FALSE)+
#scale_colour_manual(values=c(col))+
scale_x_continuous(limits=c(0,max(a2$y)), breaks=seq(from = 0, to = max(a2$y), by = 100))+
scale_y_continuous(limits=c(0,max(a2$y)), breaks=seq(from = 0, to = max(a2$y), by = 100))+
facet_wrap(~id, ncol=5)+
theme(aspect.ratio=1)+
theme(strip.text.x = element_text(size = 5))
```````
the points are already well ordered. but let's make the ordering explicit

`````{r}
a2<-a2%>%
group_by(id)%>%
mutate(pointorder = row_number())
````````
add metadata
`````{r}
b<-read.csv("Modern_pots_09122019.csv")
`````

`````{r}
a3<-merge(a2, b, by.x="id", by.y="id")
write.csv(a3, "mantamados_normandata_closed_raw.csv", row.names=FALSE)
`````



find the centre of each vase
`````{r}
a3<-a3%>%
group_by(id)%>%
mutate(minx=min(x), maxx=max(x), centrex=(maxx-minx)/2+minx)

left<-a3%>%
filter(x<=centrex)%>%
arrange(id, pointorder)
``````

`````{r}
ggplot(data=left, aes(x=x, y=y, group=id, fill=as.factor(shape_specific)))+
#geom_point(size=0.1)+
geom_polygon()+
theme_classic()+
#guides(colour=FALSE)+
#scale_colour_manual(values=c(col))+
scale_x_continuous(limits=c(0,max(a2$y)), breaks=seq(from = 0, to = max(a2$y), by = 200))+
scale_y_continuous(limits=c(0,max(a2$y)), breaks=seq(from = 0, to = max(a2$y), by = 200))+
facet_wrap(~id, ncol=5)+
theme(aspect.ratio=1)+
theme(strip.text.x = element_text(size = 5))
```````

`````{r}
write.csv(left, "mantamados_normandata_open_raw.csv", row.names=FALSE)
```````

