---
title: "Clustering Norman's standard eigenscores using PAM"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

`````{r}
rm(list=ls())
``````

`````{r, message=FALSE, warning=FALSE}
library(dplyr)
library(data.table)
library(ggplot2)
library(ape)
library(tidytree)
library(ggtree)
library(wesanderson)
library(mclust)
library(ggpubr)
library(TSclust)
library(cluster)
`````

#### data
Here we have Norman's standard eigenscore data for the Mantamados data. 

`````{r}
a<-read.csv("Mantamados_standard_eigenscores_Norman.csv", header=TRUE)
a$shape_specific<-ifelse(a$Group=="Small", "s", ifelse(a$Group=="Medium", "m", "l"))
`````````
PAM clustering.  Takes a distance matrix, so can be used in any context
````{r}
a1<-a[c(3:9)]
a2<-dist(a1, method="euclidean")
a3<-pam(a2,k=3, diss=TRUE, stand = FALSE)
a$PAMcluster<-a3$clustering
``````
examine clusters with labels
```{r}
check<-a%>%
select(Group, shape_specific, PAMcluster)
check
`````
evaluate cluster accuracy
`````{r}
cluster.evaluation(a$Group,a$PAMcluster)
`````
plot
`````{r}
a<-a%>%
arrange(Group, PC.1)
symbols<-c("c1", "c2", "c3")
pal<- wes_palette("Zissou1", 3, type = "continuous")
ggplot(a, aes(x=PC.1, y=PC.2, colour=as.factor(PAMcluster), fill=as.factor(PAMcluster), label=shape_specific))+
stat_chull(alpha = 0.5, geom = "polygon", colour=NA)+
geom_point(size=5)+
geom_text(colour="white", size=3)+
scale_colour_manual(values=pal)+
scale_fill_manual(values=pal)+
xlab("PC1")+
ylab("PC2")+
scale_y_continuous(limits=c(-4,4), breaks=c(seq(from=-4, to=4, by=1)))+
scale_x_continuous(limits=c(-4,4), breaks=c(seq(from=-4, to=4, by=1)))+
theme_classic()+
theme(aspect.ratio = 1)
`````
