---
title: "Plotting the Mantamados vase outlines"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

`````{r}
rm(list=ls())
``````

`````{r, message=FALSE, warning=FALSE}
library(dplyr)
library(data.table)
library(ggplot2)
`````

#### data
Here we have Norman MacLeod's data on the mantamados pots. We have scaled, centred and lightly smoothed the data. Now we are going to make a pretty plot

`````{r}
a<-read.csv("mantamados_normandata_closed_scaled_centred_smoothed.csv", header=TRUE)
b<-read.csv("Modern_pots_09122019.csv")
b<-b%>%
  select("id", "shape_specific")
a1<-merge(a,b, by="id")
`````````

plot to check what they look like
```{r}
library(wesanderson)
pal <- wes_palette("Cavalcanti1", 3, type = "continuous")
ggplot()+
#geom_point(data=a1, aes(x=x, y=y, colour=as.factor(shape_specific)), alpha=1, size=0.1)+  
geom_path(data=a1, aes(x=x, y=y, colour=as.factor(shape_specific)), alpha=0.5, size=0.5)+
 theme_classic()+
scale_color_manual(values = c(pal))+
scale_fill_manual(values = c(pal))+
scale_x_continuous(limits=c(-2,2), breaks=seq(from=-2, to=2, by=0.5))+
scale_y_continuous(limits=c(-2,2), breaks=seq(from=-2, to=2, by=0.5))+
theme(aspect.ratio = 1)
`````
