---
title: "Among-series Fourier d distance"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

This analyses the closed contours of the RW simulated vases using Fourier analysis in Momocs
`````{r}
library(Momocs)
library(data.table)
library(ggplot2)
library(dplyr)
library(tidyr)
```````
### data
read data 
``````{r}
long<-fread("PC_RW_large_Mantamados_closed.csv")
`````
since we are going to look at among run variance, remove the start vase
`````{r}
long<-long%>%
  filter(run!=0)
```````
look at the profile of one vase
```{r}
check<-long%>%
  filter(timestep==1000, run==1)
ggplot()+
geom_path(data=check, aes(x=x, y=y))+
xlim(-2,2)+
ylim(-2,2)+
  theme_classic()+
  theme(aspect.ratio=1)
`````
### Momocs analysis

Put in Momocs format. Split the dataframe into matrixes of individual vases.
````{r}
long<-long[,c("x","y","timestep","run")]
a<-split(long,list(long$timestep,long$run),drop=TRUE)
a<-lapply(a,function(x){as.matrix(x[,c(1,2)])})
a<-Out(a) #here we use Out since we have closed outlines of full vases
````
Centre and align the curves

coo_center returns a shape centered on the origin.
````{r}
a1<-a%>%coo_center
`````
Aligns the coordinates along their longer axis using var-cov matrix and eigen values. There are other aligning options: coo_aligncalliper, coo_alignminradius, coo_align
`````{r}
a2<-a1%>%coo_align
```````
examine what this does
````{r}
c1_a<-as.data.frame(a[1])
c1_a$transform<-"1raw"
c1_a1<-as.data.frame(a1[1])
c1_a1$transform<-"2centred"
c1_a2<-as.data.frame(a2[1])
names(c1_a2)<-c("x","y")
c1_a2$transform<-"3aligned"
c1<-rbind(c1_a, c1_a1, c1_a2)
c1$vase<-"vase1"
c500_a<-as.data.frame(a[500])
c500_a$transform<-"1raw"
c500_a1<-as.data.frame(a1[500])
c500_a1$transform<-"2centred"
c500_a2<-as.data.frame(a2[500])
names(c500_a2)<-c("x","y")
c500_a2$transform<-"3aligned"
c500<-rbind(c500_a, c500_a1, c500_a2)
c500$vase<-"vase500"
c<-rbind(c1, c500)
`````

```{r}
ggplot()+
geom_path(data=c, aes(x=x, y=y, colour=as.factor(vase)))+
facet_wrap(~transform)+
xlim(-2,2)+
ylim(-2,2)+
  theme_classic()+
  theme(aspect.ratio=1)
`````
The vases were already centred on zero.We will use the raw data.

Fourier analysis. This calculates elliptical fourier transforms. Suitable for closed curves. We extract the first 10 harmonics.
````{r}
a.f<-efourier(a,nb.h = 10, norm=FALSE)  # 10 harmonics d-fourier
``````
There are A, B, C and D coefficients. Do a PCA on them all. 
``````{r}
a.p<-as_df(a.f)
a.pca<-prcomp(a.p, scale=TRUE)
res.pca<-a.pca$x
res.pca<-as.data.frame(res.pca)
````
get the distances among vases in the same series based on these PCs
get the row labels
````{r}
wide<-long%>%
select("timestep", "run")
wide<-as.data.frame(unique(wide))
res.pca$run<-wide$run
res.pca$timestep<-wide$timestep
res.pca$virtualvase<-paste(res.pca$run, res.pca$timestep, sep="_")
`````
a function to get the distance matrix for 10 PCs
`````{r}
distancematrix<- function(x){
d1<-x[,c(1:40)]# change matrix size
d2<-as.matrix(d1)
d3<-dist(d2, method="euclidean")
d4<-as.matrix(d3)
d5<-as.data.frame(d4)
return(d5)
}
`````
get mean distance by timestep
`````{r}
timestep_dist<-res.pca%>%
group_by(timestep)%>%
do(d1=distancematrix(.))%>%
summarize(q50=quantile(as.matrix(d1), 0.5), q05=quantile(as.matrix(d1), 0.05), q95=quantile(as.matrix(d1), 0.95))
timestep_dist$timestep<-1:1000
`````
check that the distance matrix estimation function is working correctly
`````{r}
check<-timestepdist<-res.pca%>%
filter(timestep==1000)%>%
select(PC1:PC40)
distance<-dist(as.matrix(check), method="euclidean")
distance<-as.matrix(distance)
q50<-quantile(distance,0.5)
q50
tail(timestep_dist)
```````
it works
````{r}
m1<-lm(q50~0+timestep, data=timestep_dist)
timestep_dist$fit<-predict(m1)
summary(m1)
#plot(m1)
``````
plot
````{r}
ggplot()+
geom_ribbon(data=timestep_dist, aes(ymin=q05, ymax=q95, x=timestep), fill="grey50", alpha=0.5)+
geom_point(data=timestep_dist, aes(x=timestep, y=q50), colour="grey50", size=1)+
guides(colour=FALSE)+
geom_line(data=timestep_dist, aes(x=timestep, y=fit))+
labs(x = "time step", y = "distance")+
theme_classic()+
theme(aspect.ratio=1)
```````
