---
title: "Among-series Fourier d distance"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

This analyses the open contours of the RW simulated vases using Fourier analysis in Momocs
`````{r}
library(Momocs)
library(data.table)
library(ggplot2)
library(dplyr)
library(tidyr)
```````
### data
read data 
``````{r}
long<-fread("PC_RW_large_Mantamados_open.csv")
`````
since we are going to look at among run variance, remove the start vase
`````{r}
long<-long%>%
  filter(run!=0)
```````
look at the profile of one vase
```{r}
check<-long%>%
  filter(timestep==1000, run==1)
ggplot()+
geom_path(data=check, aes(x=x, y=y))+
xlim(-2,2)+
ylim(-2,2)+
  theme_classic()+
  theme(aspect.ratio=1)
`````
### pre processing
to make the open contours most comparable to the closed contours, we want to have the top and the bottom of the virtual vases reach to the maximum x value (so that they are genuinely half a vase). This is always -0.01
````{r}
long1<-long%>%
group_by(run, timestep)%>%
summarize(maxx=max(x))
head(long1)
``````
the top and bottom y values are
````{r}
long2<-long%>%
group_by(run, timestep)%>%
summarize(maxy=max(y), miny=min(y))
head(long2)
`````
adjust
````{r}
long$x<-ifelse(long$y==-1.339148| long$y==1.418529, -0.01, long$x)
``````
check
```{r}
check<-long%>%
  filter(timestep==1000, run==1)
ggplot()+
geom_path(data=check, aes(x=x, y=y))+
xlim(-2,2)+
ylim(-2,2)+
  theme_classic()+
  theme(aspect.ratio=1)
`````
### Momocs analysis

Put in Momocs format. Split the dataframe into matrixes of individual vases.
````{r}
long<-long[,c("x","y","timestep","run")]
a<-split(long,list(long$timestep,long$run),drop=TRUE)
a<-lapply(a,function(x){as.matrix(x[,c(1,2)])})
a<-Opn(a) #here we use Opn since we have open outlines of half vases
````
Centre and align the curves

coo_center returns a shape centered on the origin.
````{r}
a1<-a%>%coo_center
`````
Aligns the coordinates along their longer axis using var-cov matrix and eigen values. There are other aligning options: coo_aligncalliper, coo_alignminradius, coo_align
`````{r}
a2<-a1%>%coo_align
```````
examine what this does
````{r}
c1_a<-as.data.frame(a[1])
c1_a$transform<-"1raw"
c1_a1<-as.data.frame(a1[1])
c1_a1$transform<-"2centred"
c1_a2<-as.data.frame(a2[1])
names(c1_a2)<-c("x","y")
c1_a2$transform<-"3aligned"
c1<-rbind(c1_a, c1_a1, c1_a2)
c1$vase<-"vase1"
c500_a<-as.data.frame(a[500])
c500_a$transform<-"1raw"
c500_a1<-as.data.frame(a1[500])
c500_a1$transform<-"2centred"
c500_a2<-as.data.frame(a2[500])
names(c500_a2)<-c("x","y")
c500_a2$transform<-"3aligned"
c500<-rbind(c500_a, c500_a1, c500_a2)
c500$vase<-"vase500"
c<-rbind(c1, c500)
`````

```{r}
ggplot()+
geom_path(data=c, aes(x=x, y=y, colour=as.factor(vase)))+
facet_wrap(~transform)+
xlim(-2,2)+
ylim(-2,2)+
  theme_classic()+
  theme(aspect.ratio=1)
`````
The vases were already aligned on their right. I don't think that they need any more alignment, or if they do, they only need centering.  We will use the raw data.

Fourier analysis. This calculates discrete cosine fourier transforms. Suitable for open curves. We extract the first 10 harmonics.
````{r}
a.f<-dfourier(a,nb.h = 10)  # 10 harmonics d-fourier
``````
There are A and B harmonic coeffients. Which should be used? Both? The amplitudes of the cosine waves are held in the variables: a1, a2, a3, a3, etc., while the amplitudes of the sine waves are held in: b1, b2, b3, b4, and so on. In other words, the "a" and "b" coefficients are the real and imaginary parts of the frequency spectrum, respectively. I think use them all.

do a PCA on just the a coefficients
``````{r}
a.p<-as_df(a.f)
a.pca<- prcomp(a.p, scale=TRUE)
res.pca<-a.pca$x
res.pca<-as.data.frame(res.pca)
```````
get the row labels
````{r}
wide<-pivot_wider(long,names_from = y, values_from=x)
res.pca$run<-wide$run
res.pca$timestep<-wide$timestep
res.pca$virtualvase<-paste(res.pca$run, res.pca$timestep, sep="_")
`````
a function to get the distance matrix for X PCS
`````{r}
distancematrix<- function(x){
d1<-x[,c(1:20)]
d2<-as.matrix(d1)
d3<-dist(d2, method="euclidean")
d4<-as.matrix(d3)
d5<-as.data.frame(d4)
return(d5)
}
`````
get mean distance by timestep
`````{r}
timestep_dist<-res.pca%>%
group_by(timestep)%>%
do(d1=distancematrix(.))%>%
summarize(q50=quantile(as.matrix(d1), 0.5), q05=quantile(as.matrix(d1), 0.05), q95=quantile(as.matrix(d1), 0.95))
timestep_dist$timestep<-1:1000
`````
check that the distance matrix estimation function is working correctly
`````{r}
check<-timestepdist<-res.pca%>%
filter(timestep==1000)%>%
select(PC1:PC20)
distance<-dist(as.matrix(check), method="euclidean")
distance<-as.matrix(distance)
q50<-quantile(distance,0.5)
q50
tail(timestep_dist)
```````
it works
````{r}
m1<-lm(q50~0+timestep, data=timestep_dist)
timestep_dist$fit<-predict(m1)
summary(m1)
#plot(m1)
``````
plot
````{r}
ggplot()+
geom_ribbon(data=timestep_dist, aes(ymin=q05, ymax=q95, x=timestep), fill="grey50", alpha=0.5)+
geom_point(data=timestep_dist, aes(x=timestep, y=q50), colour="grey50", size=1)+
guides(colour=FALSE)+
geom_line(data=timestep_dist, aes(x=timestep, y=fit))+
labs(x = "time step", y = "distance")+
  theme_classic()+
  theme(aspect.ratio=1)
```````
