---
title: "Among-series variance of RW vase eigenscores"
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

Estimating the variance of individual variables, by interval of RW vases. Here we are going to estimate the variance of each variable and series one at a time. We estimate the variance of the variable at time point x relative to start.
`````{r}
library(Momocs)
library(data.table)
library(ggplot2)
library(dplyr)
library(reshape2)
library(matrixStats)
library(tidyr)
library(broom)
```````
This estimates the variance among runs for each eigen value

read data and put in Momocs format
``````{r}
wide<-fread("PC_RW_large_Mantamados_eigenscores.csv")
````````
reshape data

````{r}
l<-wide%>%
pivot_longer(col=c("1":"4"), names_to="PC", values_to="eigenvalue")
l1<-l%>%
  pivot_wider(id_cols=c(timestep, PC), names_from=run, values_from=eigenvalue)%>%
  arrange(PC, timestep)
`````

```{r}
l2<-as.matrix(as.data.frame(l1[c(3:102)]))
l1$variance<-rowVars(l2, dim=c(4000, 100), na.rm=TRUE)
l3<-l1%>%
  select(timestep, PC, variance)
``````
plot
````{r}
ggplot()+
geom_point(data=l3, aes(x=timestep, y=variance, colour=as.factor(PC)), size=0.5, alpha=0.5)+
#guides(colour=FALSE)+
geom_smooth(data=l3, aes(x=timestep, y=variance, colour=as.factor(PC)), se=FALSE, method="lm", formula=y~0+x, size=0.5)+
theme_classic()+
theme(aspect.ratio=1)
```````
do regressions on them all
`````{r}
m1<-l3 %>% group_by(PC) %>%
  do(fit = lm(variance ~ 0+timestep, data = .))
coefs<-tidy(m1, fit)%>%
filter(term=="timestep")
coefs
`````
rsq
````{r}
rsq<-glance(m1, fit)
``````
get the average rsquared
````{r}
mean(rsq$adj.r.squared)
```````
get the predictions
`````{r}
pred<-augment(m1, fit)%>%
select(PC, timestep, .fitted)%>%
  arrange(PC, timestep)
```````
order for plotting
`````{r}
o<-coefs%>%
select(PC,estimate)%>%
arrange(desc(estimate))
o$order<-1:nrow(o)
l3<-merge(l3, o, by="PC")
pred<-merge(pred, o, by="PC")
l3$PC<- reorder(l3$PC, l3$order)
pred$PC<- reorder(pred$PC, pred$order)
levels(pred$PC)
`````

````{r}
library(wesanderson)
pal <- wes_palette("GrandBudapest1", 4, type = "continuous")
ggplot()+
geom_point(data=l3, aes(x=timestep, y=variance, colour=as.factor(PC)), size=0.5, alpha=0.5)+
#guides(colour=FALSE)+
geom_line(data=pred, aes(x=timestep, y=.fitted, colour=as.factor(PC)))+
scale_color_manual(values = c(pal))+
scale_fill_manual(values = c(pal))+
scale_x_continuous(breaks=seq(from=0, to=1000, by=250))+
scale_y_continuous(breaks=seq(from=0, to=14, by=2))+
theme_classic()+
theme(aspect.ratio=1)
```````  
